---
title: "523 Final Project"
author: "Outliers: Ryan Fang, Qinzhe Wang, Congwei Yang, Haoliang Zheng"
date: "November 22, 2020"
output: 
  html_document:
    theme: readable
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, comment = FALSE, 
                      message = FALSE, warning = FALSE, fig.align = "center")
```

# Introduction
***
Our project goal is to retrieve most up-to-date football match data from online 
API and use a shiny app to display, visualize, and model the data. 
`API-Football` contains the real-time data for most major leagues worldwide. In 
this project, we will use the data from `API-Football` and focus on three major 
European leagues: Premier League of England, Primera Division (La Liga) of Spain, 
and Bundesliga of Germany. We will construct multiple API wrapper functions based 
on the API structure to help us scrape data in the Shiny App, and use the data 
scraped to build visualization and clustering model within the App. Moreover, 
the App will also serve as a football match info center that provides team/play 
information search, player/league statistics summary and real-time match schedule 
panel. 


# API wrapper functions
***
We constructed mutiple API wrapper function to use in the Shiny App. They have 
similar structure but different parameters and returned values based on the 
structure of the API. The following code chunk is a example wrapper function. 

```{r wrapper-function1}
get_player_by_team_season <- function(
  # apikey
  apiKey = "65563b231bmsh5e04da3f6a146c5p1b1625jsn0e31c1d79a15",
  # parmaters
  team = 33, 
  season = "2018-2019"){
  
  
  url_player <- paste0(
    "https://api-football-v1.p.rapidapi.com/v2/players/team/", team, "/", season)
  response_player_team <- VERB("GET", 
                          url_player, 
                          add_headers('x-rapidapi-host'='api-football-v1.p.rapidapi.com',
                                      'x-rapidapi-key'= apiKey),
                          content_type("application/octet-stream"))
  raw_player_team <- content(response_player_team, "text",encoding = "UTF-8")
  raw_table_player_team <- tibble(fromJSON(raw_player_team)$api$players)
  table_player_team = flatten(raw_table_player_team)
  return(table_player_team)
}
```

The `API-Football` uses JSON format for all endpoints. Thus, for every wrapper 
function, after the API call, we use `fromJSON()` and `tibble()` or 
`as.data.frame()` to convert the data into workable format. 

The following is a list of our wrapper functions: 

1. `get_last_fixture()`: Get a number of recent finished matches in the specified 
league and their corresponding data. The function takes `league_id` as parameter. 

2. `get_next_fixtures()`: Get a number of future upcoming matches in the specified 
league and their corresponding information. The function takes `league_id` as 
parameter. 

3. `get_league()`: Search all available leagues in the `API-Football`. 

4. `get_player()`: Retrieve the data of a player by specifying `player_id`. 

5. `get_player_by_team_season()`: Retrieve the data for all players of a team in 
a season by specifying the `season` and `team_id`. 

6. `get_team()`: Retrieve team information for all teams in the specified league. 
The function takes `league_id` as parameter. 

7. `get_team_data()`: Retrieve match data for specified team by input `league_id` 
and `team_id`. 

8. `get_team_fixture()`: Retrieve data for each game that the specified team 
participated by input `league_id` and `team_id`. 

9. `get_top_scorers()`: Get the player data for top 20 scorers in the specified 
league by input `league_id`. 

10. `get_standings()`: Get the final ranking and cumulative game results for 
each team in the specified league by input `league_id`. 

A noticeable property of the commonly used parameter `league_id` is that due to 
the fact the teams in a league may change across different years, the `league_id` 
will be different for the same league in different years. For example, Premier 
League in 2018 and Premier League in 2019 have different `league_id`. We also 
constructed a league-year-ID matching table for the leagues we are working with. 

# Clustering function
***
Clustering is a technique to find similar observations in a dataset mostly based 
on the Euclidean distances among those observations. In this way, the dataset is
partitioned into K distinct and non-overlapping subsets. This is very useful for 
pattern recognizing and potential classification. Since K-means clustering does 
not require the labeling of the observations while the model is trained,it is an 
unsupervised machine learning algorithm in clustering observations. Therefore, 
this algorithm is most attractive by its effectiveness, easy handling, and 
adaptiveness to a variety of unlabeled datasets. Here is an example of the 
Algorithm of the K-means clustering.  

1. Randomly assign a number, from 1 to K, to each of the observations. These serve 
as initial cluster assignments for the observations.  
2. Iterate until the cluster assignments stop changing:
-item For each of the K clusters, compute the cluster centroid. The k-th cluster 
centroid is the vector of the p feature means for the observations in the k-th 
clustere.  
-item Assign each observation to the cluster whos centroid is closest (where 
closest is defined using Euclidean Distances.)  


We construct this K-means clustering algorithm to group the players and teams 
given certain football match seasons and leagues with a hyperparameter k. And we 
also create animations showing how the players and teams are graduated clustered 
using this algorithm. By showing the clustering result of the player and team, 
people might discover some pattern among the players and teams which are not 
revealed before. We use the `assist_per_game`, `goals_per_shot`, `ry_card_per_game`, 
`goals_per_hour`, and `goals_per_shot` to characterize and cluster the players. 
We use the `win`, `draw`, `lose`, `goals`, and other 11 matching related features 
to character and cluster the teams. We run Principle Component Analysis (PCA) 
on the features to reduce the number of features to two that they could be 
displayed and animated on a two-dimensional plot. Here is a list of functions 
developed for K-means clustering and visualization.  


1. `show_cluster()`: show cluster results with varied K centroids on th `USArrest` 
dataset for a simple demonstration of K-means clustering algorithm.  

2. `get_data()`: retrieve "player" and "team" data given the certain league and 
season. It also could generate data randomly. The principle component analysis 
and data transformatino is done in this function if necessary.  

3. `myKeans()`: this function implements K-means clustering algorithm on the
dataset. It takes a list of parameters including, the number of centroids (k),
the stopping criteria etc. It also returns a list of variables characterizing 
the K-means clustering result such as, cluster result, cluster history, centroid
history, with cluster sum of squares, and etc.  

4. `cluster()`: this is the function wraps the `myKmeans()` function to prepare the 
data for animation in plots and summary in tables.  

5. `animation()`: this function complete the animation of the data, in which the 
K-means clustering process is shown step by step so that the reader could easily 
understand how the results are obtained.  

To implement this algorithm, we also use our functions on our APP. As for the
performance on the football data on our APP. The clustering results sometimes
could be unstable because the number of the observations are limited to 20. The
last four functions shown above serve as a modeling and visualization pipeline,
which is adapted to "player", "team" and "random" data. The visualization is a 
step-by-step clustering and assignment process, shown in a two-dimensional plot
with clusters labeled by different colors. The metadata results related to the
clusters are also summarized in the tables for the readers to explore.  

In the clustering page, there are two tabs, one for the team clustering and the 
other for scorer clustering. On the team clustering tab, after selecting the 
season and the number of clusters, the visualization and summary tables are 
displayed. The same procedure also works on the scorer cluster tab. There are
also clustering demo displayed in the `More` page, one for the USArrest crime
dataset clustering and one for the clustering on a randomly generated dataset.
USArrest crime dataset clustering is simply to show what kind of results are
generated from the clustering algorithm. The randomly generated data contains
more than three hundred observations, so the K-means clustering result is stable
and the clustering process is shown to the readers step by step to see how the
clustering algorithm work in an interpretable and direct way.    

# Shiny APP
***

Our Shiny App consists of 4 tabs: Recent fixture page that also serves as a 
real-time match schedule panel, a Team/Player information search page, a 
Team/Top Scorers data summary/visualization page, and a clustering model page. 

1. The Recent Fixture Page. 

When the user first open up the App, there will be a modal dialogue window pop 
up and the user will be asked to select one of the three leagues: Premier League, 
Primera Division (La Liga), and Bundesliga. After the user made the selection, 
all content that the app display will be based on the selected league. the 
recent fixture page will display (ordering from top to bottom) currently ongoing 
match ("No fixtures for now" will be showed if there is no ongoing match), most 
recent 3 finished matches (the user can scroll down to look at up to most recent
10 matches), and the upcoming 3 matches (the user can also scroll down to look 
at up 10 upcoming matches). The user can also see the league logo on the 
top-right corner of the page and the team logo for each team in the schedule 
chart. 

2. Team/Player information search page. 

The second tab is the Team/Player information search page. There are two sub-tabs, 
one for team information search and another for player information search. 
For the team information search, the user need to select a `Season` and click 
on the `Search Teams`, then the app will display a table containing information 
of all teams played during that season in the selected league. For the player 
search page, the user need to select a `Season` and click on `Select This Season`. 
Then a new input widget will pop up and user can use the drop-down menu to select
a team that has played during that season in the specified league. Finally, the 
user can click on the button `Search Players`. The App will display the team logo
under the input widget section, and on the right side a table of player information
will be displayed. 

3. Team/Top Scorers data summary/visualization. 

The third tab is the Team/Top Scorers data summary and visualization. Similar 
with the structure of last tab, there are two sub-tabs, one for team data and 
another for top scorers. 

For the team data summary/visualization page, the user need to select a `Season`
and click `Select This Season`. Then a new input widget will appear on the right 
side and the user can use the drop-down menu to select a team. After selection, 
the user can click on `Show Team Record` and the App will generate donut charts 
for the team records in the selected season. The App will also show a win/lose 
path of the selected team in that season. 

In top scorers page, the user need to select a `Season` and click 
`Select This Season`. Then a new widget will appear and the user can select the 
player to visualize. The first graph is a word cloud. The "frequency" is based 
on the total number of goals of the selected player during this season. The 
second graph is radar chart for the player selected. The user can evaluate the 
player based on five dimensions: Goals per Game, Assist per Game, Goals per Shot, 
Red/Yellow Card per Game, Goals per Hour. 

4. "Clustering" Panel

We have written our own functions for clustering algorithm. We implement them 
here on football data, and users can try different number of clusters by 
themselves. There are also both team-level analysis and player-level analysis. 
You can get the result from table also with visualizations below. The running 
time may be a little bit long because we're making animation. Please wait for 
a second!

# Reference

First, thanks to the [football API](https://www.api-football.com/) where we get
data from.

Seconds, there are few examples and Q&As that inspire us to achieve some 
functions in my website, I will list them below:

- https://gallery.shinyapps.io/010-download/

  From this website, we learned how to achieve the download function.

- https://shiny.rstudio.com/gallery/word-cloud.html
  
  From this website, we learned how to draw word cloud.

- https://www.r-graph-gallery.com/spider-or-radar-chart.html
  
  From this website, we learned how to draw radar chart.
  
- https://datatables.net/reference/option/
  
  From this website, we learned how to set detailed options for the datatable.
  
- https://www.r-graph-gallery.com/128-ring-or-donut-plot.html

  From this website, we learned how to make donut/ring charts using ggplot. 

- https://shiny.rstudio.com/gallery/word-cloud.html
  
  From this website, we learned how to draw word cloud.

- https://www.r-graph-gallery.com/spider-or-radar-chart.html
  
  From this website, we learned how to draw radar chart.
  
- https://datatables.net/reference/option/
  
  From this website, we learned how to set detailed options for the datatable. 

- https://smorbieu.gitlab.io/animate-intermediate-results-of-your-algorithm/
  
  From this website, we learned how to implement some K-means algorithm in details 
  some related animation in R. 

- Gareth James, Daniela Witten, Trevor Hastie, Robert Tibshirani. An Introduction 
to Statistical Learning : with Applications in R. New York :Springer, 2013.  
  
  From this book, we learned how the K-means algorithm works and how to implement 
  the K-means clustering algorithm in R.  

- https://www.datanovia.com/en/blog/gganimate-how-to-create-plots-with-beautifu
l-animation-in-r/  
  
  From this website, we learned how to create animations.  

- https://stackoverflow.com/questions/35421923/how-to-create-and-display-an-anim
ated-gif-in-shiny  
  
  From this website, we learned how to implete the animation in the R-shiny app.

